A performance comparison of Dask and Spark for data-intensive neuroimaging pipelines

### 2019-07-29 latex/paper.tex
Reduce wording

%Similarities
Spark and Dask both offer in-memory computing, data locality, and lazy evaluation;
which is common for Big Data frameworks. Both their schedulers operate dynamically,
which is good when task runtimes are not known ahead of time~\cite{Dask:15}. They
also provide rich, high-level programming APIs, and support a variety of
infrastructure schedulers, such as Mesos~\cite{hindman2011mesos}, YARN~\cite{vavilapalli2013apache}, or HPC clusters (Dask only). Over
these similarities, the frameworks have differences.

% Differences
First and foremost, Spark is written in Scala while Dask is in Python.
Given the popularity of Python in scientific communities, this arguably
gives an edge to Dask due to data serialization costs from Python to Scala.
On the other hand, Python's Global Interpreter Lock (GIL) might reduce
parallelism in some cases. This difference in programming languages also
has qualitative implications. As part of the \href{http://scipy.org}{SciPy}
ecosystem, Dask provides almost transparent parallelization of applications
manipulating Numpy arrays or Pandas data frames. On the other hand, Spark's
Java, R and Python APIs allow to easily parallelize analyses that combine
these languages, with reduced performance loss.

% Related work
Spark and Dask were both included in the evaluation reported
in~\cite{Mehta:17} that used a neuroimaging and an astronomy application
processing approximately \SI{100}{\giga\byte} of data each. In this work, Dask was
reported to have a slight performance advantage over Spark. Overall, Dask's
end-to-end time (makespan) was measured to be up to 14\% faster than Spark,
due to ``more efficient pipelining'' %(Fig 10) 
and serialization time to
Python. % (Fig. 12-a)
Dask, however, was reported to have a larger startup
time than Spark. % (Fig. 12-c)
The analysis remains at a quite high level
though, leaving most of the observed performance difference unexplained. In
comparison, our study will provide a detailed analysis of performance
differences and similarities between Spark and Dask.

@@

\subsection{Effect of tasks' size}
We measured the effect of various block sizes for constant BigBrain image for
Incrementation, Histogram, and Histogram (NumPy). We observe that larger block size
results in more IO compensated by lower overhead time. On the other hand, smaller
block size results in higher overhead time compensated by reduced IO time.

@@

Varying the block size is equivalent to varying the size of the partitions. We
observed no difference when varying the task size while using an NFS. However, if
one uses a parallel shared file system, we suppose it would be preferable to use a
larger block size as the IO time should be reduced substantially. It is worth noting
that the Dask developers suggest, in
\href{https://docs.dask.org/en/latest/best-practices.html}{Dask's best practices}, to
use a partition size such that workers have 2-3 times as many chunks available to
work on. More experiments would have to be performed to gain insight on an optimal
number of partitions to use for a given set of data-intensive applications.

>>>>

Spark and Dask both offer in-memory computing, data locality, and lazy evaluation;
which is common for Big Data frameworks. Both their schedulers operate dynamically,
which is good when task runtimes are not known ahead of time~\cite{Dask:15}. They
also provide rich, high-level programming APIs, and support a variety of
infrastructure schedulers, such as Mesos~\cite{hindman2011mesos}, YARN~\cite{vavilapalli2013apache}, or HPC clusters (Dask only). Over
these similarities, the frameworks have differences.

First and foremost, Spark is written in Scala while Dask is in Python.
Given the popularity of Python in scientific communities, this arguably
gives an edge to Dask due to data serialization costs from Python to Scala.
On the other hand, Python's Global Interpreter Lock (GIL) might reduce
parallelism in some cases. This difference in programming languages also
has qualitative implications. As part of the \href{http://scipy.org}{SciPy}
ecosystem, Dask provides almost transparent parallelization of applications
manipulating Numpy arrays or Pandas data frames. On the other hand, Spark's
Java, R and Python APIs allow to easily parallelize analyses that combine
these languages, with reduced performance loss.

Spark and Dask were both included in the evaluation reported
in~\cite{Mehta:17} that used a neuroimaging and an astronomy application
processing approximately \SI{100}{\giga\byte} of data each. In this work, Dask was
reported to have a slight performance advantage over Spark. Overall, Dask's
end-to-end time (makespan) was measured to be up to 14\% faster than Spark,
due to ``more efficient pipelining'' %(Fig 10) 
and serialization time to
Python. % (Fig. 12-a)
Dask, however, was reported to have a larger startup
time than Spark. % (Fig. 12-c)
The analysis remains at a quite high level
though, leaving most of the observed performance difference unexplained. In
comparison, our study will provide a detailed analysis of performance
differences and similarities between Spark and Dask.

@@

\subsection{Effect of tasks' size}
We measured the effect of various block sizes for constant BigBrain image for
Incrementation, Histogram, and Histogram (NumPy). We observe that larger block size
results in more IO compensated by lower overhead time and vice-versa.

@@

Varying the block size or the size of the partitions is equivalent. No difference is
observed when varying the task size while using an NFS. However, if one uses a
parallel shared file system, we suppose it would be preferable to use a larger block
size as the IO time should be reduced substantially. It is worth noting that the Dask
developers suggest, in
\href{https://docs.dask.org/en/latest/best-practices.html}{Dask's best practices}, to
use a partition size such that workers have 2-3 times as many chunks available to
work on. More experiments would have to be performed to gain insight on an optimal
number of partitions to use for a given set of data-intensive applications.
